State and Input Dependent Disturbances in Min-Max Model Predictive Control

### 2015-10-23 draft.tex
First draft.
what happened

[ticket: X]

@@

\newcounter{thmcount}
\renewcommand{\thethmcount}{\arabic{thmcount}}
\renewcommand{\theequation}{\arabic{section}.\arabic{equation}}
\renewcommand{\thefigure}{\arabic{section}.\Roman{figure}}

@@

\section{Introduction}
% The very first letter is a 2 line initial drop letter followed
% by the rest of the first word in caps.
% 
% form to use if the first word consists of a single letter:
% \IEEEPARstart{A}{demo} file is ....
% 
% form to use if you need the single drop letter followed by
% normal text (unknown if ever used by IEEE):
% \IEEEPARstart{A}{}demo file is ....
% 
% Some journals put the first two words in caps:
% \IEEEPARstart{T}{his demo} file is ....
% 
% Here we have the typical use of a "T" for an initial drop letter
% and "HIS" in caps to complete the first word.
\IEEEPARstart{T}{hroughout} this paper we denote by $f\vert_{\mathcal C} = \{y:\exists x\in\mathcal C\;y=f(x)\}$
% You must have at least 2 lines in the paragraph with the drop letter
% (should never be an issue)

@@

  \mathcal W(p) := \left\{x: a_i x \leq \max_{k\leq l}\{b_{i,k} + C_{i,k}p\},i\leq m\right\}

@@

\item Notice that $h_i(p) = \max_{k\leq l} \{b_{i,k} + C_{i,k}p\}$ is a multi-parametric linear program (mpLP),
its solution is a piecewise affine function $h_i(p) = b_{i,k^\ast_i} + C_{i,k^\ast_i}p$ where $k^\ast_i(p)$ is constant

@@

  \mathscr G(\mathcal W) = \bigcup_{j\leq t} \left\{x: a_i x \leq b_{i,k_i^\ast} + C_{i,k_i^\ast}p,i\leq m

@@

$\left(b_{i,k_i^\ast} + C_{i,k_{j_1}^\ast} p\right)\big\vert_{\mathcal P_{j_1}} = 
\left(b_{i,k_{j_2}^\ast} + C_{i,k_{j_2}^\ast} p\right)\big\vert_{\mathcal P_{j_2}}$.

@@

  a_i v_\kappa(p) &= b_{i,k_i^\ast} + C_{i,k_i^\ast} p \quad\forall i\in\mathcal A_\kappa(p)\\
  a_i v_\kappa(p) &< b_{i,k_i^\ast} + C_{i,k_i^\ast} p \quad\forall i\in\mathcal I_\kappa(p)

@@

  \frac{d}{dp}\left(  A_{\mathcal J\cup \{s\}} v_{\kappa_1}(p) - b_{\mathcal J\cup \{s\}} - C_{\mathcal J\cup \{s\},{\bf{k}}^\ast} p\right) &= 0\\
  \frac{d}{dp}\left(  A_{\mathcal J\cup \{u\}} v_{\kappa_2}(p) - b_{\mathcal J\cup \{u\}} - C_{\mathcal J\cup \{u\},{\bf{k}}^\ast} p\right) &= 0

@@

  A_{\mathcal J\cup \{s\}} \frac{dv_{\kappa}}{dp} &= C_{\mathcal J\cup \{s\},{\bf{k}}^\ast}\\
  A_{\mathcal J\cup \{u\}} \frac{dv_{\kappa}}{dp} &= C_{\mathcal J\cup \{u\},{\bf{k}}^\ast}

@@

  \frac{dv_\kappa}{dp} = A_{\mathcal J\cup \{s\}}^{-1}C_{\mathcal J\cup \{s\},{\bf{k}}^\ast} = 
  A_{\mathcal J\cup \{u\}}^{-1}C_{\mathcal J\cup \{u\},{\bf{k}}^\ast}

@@

Corollary~\ref{thm:p:convexity:PWA:set:constant:num:verts} is essential for numerical examples, as it can be reformulated to:

@@

for a linear system~$x^+=\Psi x + v$ subject to disturbance
$v\in\mathcal V(x)$ where $\mathcal V(x)$ is a piecewise affine parametrically convex point-to-set map defined 

@@

Since ${\mathcal{V}}(x)$
has a piecewise affine dependence on $x$ the vertices $v_i(x)$ are also piecewise affine in $x$.
The set $\mathcal X^\infty$ as defined in~\eqref{eq:definition:MRPI:set:state:dependent} 
is required to satisfy $\Psi x + v\in\mathcal X^\infty$ for all
$x\in\mathcal X^\infty$ and $v\in\mathcal V(x)$. To compute the MRPI set we start from the given 
state constraint set

@@

recursively \emph{cut off} points that cannot satisfy the invariance condition~\eqref{eq:definition:MRPI:set:state:dependent}

@@

$\mathcal X_0$. The first iteration enforces the constraint:

@@

Here $v_{0,i}^\ast(x)$ is not necessarily given by a unique maximiser, 
but is the solution of a multi-parametric linear program and hence is given 
by a vertex $v_i(x)$ of $\mathcal V(x)$ for each $x$ on that facet.

@@

% \begin{align*}
%   v_{l,i}^\ast(x)&=\max_j \ \genmat_{0,i}\Psi^{l-1}v_j(x)
%    = \begin{array}[t]{rl} \displaystyle\max_v & \genmat_{0,i}\Psi^{l-1}v \\ \text{s.t.}& v\in\mathcal V(x)
%    \end{array} \\ 
%    &= \begin{array}[t]{rl} \displaystyle\max_{\tilde v} & \genmat_{0,i}\tilde v\\ \text{s.t.}& \tilde v\in 
%    \Psi^{l-1}\mathcal V(x) \end{array}
% \end{align*}

@@

Determining the map $v_{l,i}(x)$ for all admissible $x\in \mathcal X_0$ in general poses a great classification

@@

Notice that we do not require $\Psi$ to be invertible, in~\eqref{eq:set:iteration:state:dependent:constraints} $\Psi^{-1}\mathcal X_k$
merely denotes the preimage of $\mathcal X_k$ under the linear map $\Psi$.

@@

Let the state constraint set $\mathcal X_0$ be contained in a band: $\mathcal X_0\subseteq B=\{x:\Gamma x\leq{\bf{1}}\wedge 
-\Gamma x\leq{\bf{1}}\}$, let the pair $(\Psi,\Gamma)$ be observable and let $\mathcal V(x)$ be defined 
by~\eqref{eq:PWA:disturbance:set} in terms of a piecewise affine function $H(x)$. Then $\mathcal X_N\subseteq \mathcal 
X_{N+1}$ for a finite $N$, and hence the MRPI set $\mathcal X^\infty =\mathcal X_N$ is a polytope.

@@

if $\mathcal X_k =\emptyset$ for any $k\geq 0$. For the remainder of the proof we assume that the MRPI set 
$\mathcal X^\infty$ is non-empty. For this we require that $\bigcup_{s\in\mathcal S}\mathcal V(s)$ is 
bounded on any bounded set $\mathcal S\subset\mathbb R^n$.

The proof has two main steps: First we prove that $\mathcal X_p$ is bounded for $p\leq n$
where $n$ is the dimension of the state $x$. The second step is to prove that in~\eqref{eq:set:iteration:state:dependent:constraints}
the set $D_k$ grows exponentially, i.e. that for any given bounded set $\mathcal C$ there exists
finite $N$ such that $\mathcal C\subseteq D_{N}$. The proof is concluded by setting 
$\mathcal C = \mathcal X_{N}$ and deducing that $\mathcal X^\infty = \mathcal X_N$.

For the first step, notice that observability of $(\Psi,\Gamma)$ is equivalent to the observability matrix 
$\Omega$ having full rank, i.e.\ $\mathrm{rank}(\Omega) = \mathrm{rank}(\begin{bmatrix} \Gamma^T \ \cdots \ 
(\Gamma\Psi^{n-1})^T\end{bmatrix}) = n$.
But  this implies that the set 
%
\[
\mathcal P_{n} = \{x: 
\Omega x\leq{\bf{1}}\wedge-\Omega x\leq{\bf{1}}\} = \bigcap_{0\leq l\leq n-1} \Psi^{-l} B
\]
%
is bounded, and since $\mathcal X_k\subseteq \mathcal P_k$ for all $k$, 
the set $\mathcal X_{n}$ is also bounded. To show that $D_k$ grows exponentially we use the boundedness 
of $\mathcal V(\mathcal X_{n})$. Let $K_1$ denote the smallest ball that contains $\mathcal V(\mathcal X_{n})$ and let
$\rho$ denote the spectral radius of $\Psi$. Furthermore, let $K_2$
be the biggest ball that is contained in $\mathcal X_{0}$. From~\eqref{eq:set:iteration:state:dependent:constraints}, 
$D_k$ has the representation
%
\begin{equation}
D_k = \underbrace{\Psi^{-k}\mathcal X_0}_{=\mathcal S_1} \ominus \underbrace{\biggl(\bigoplus_{1\leq i\leq k-1} 
\Psi^i\mathcal V(\mathcal X_k)\biggr)}_{=\mathcal S_2}.
\end{equation}
%
Since $\Psi$ is asymptotically stable we have $\rho<1$ and therefore $\rho^{-k}K_2\subseteq\Psi^{-k} K_2 
\subseteq \Psi^{-k}\mathcal X_0 = \mathcal S_1$, implying exponential growth of $D_k$ provided $\mathcal S_2$ 
is bounded. To bound $\mathcal S_2$ we note that for $k\geq n$ we have $\mathcal S_2\subseteq 
\bigoplus_{1\leq i\leq k-1} \rho^i K_1 \subseteq \frac{1}{1-\rho} K_1$. Thus $D_k$ grows exponentially, whereas 
$\mathcal X_k$ is contained in a ball of finite radius, and we conclude from~\eqref{eq:set:iteration:state:dependent:constraints} that 
$\mathcal X_{k+1} = \mathcal X _k$ for all $k\geq N$ and hence $\mathcal X _N = \mathcal X^\infty$ for finite $N$.
\end{proof}
%
We have seen that we can compute the maximal robust positively invariant set $\mathcal X^\infty$ for linear 
systems with state dependent constraints. In the next section we will illustrate the concept with an 
example.

\subsection{Example I}

\section{Recursive Multi-Parametric Quadratic Programming to Solve Min-Max Programs}

\section{Solving Min-Max Programs}

\section{Example II}

>>>>

\usepackage{lpic}

\usepackage{xcolor}

@@

\newcounter{thmcount}
\renewcommand{\thethmcount}{\arabic{thmcount}}
\renewcommand{\theequation}{\arabic{section}.\arabic{equation}}
\renewcommand{\thefigure}{\arabic{section}.\arabic{figure}}

@@

\section{Introduction}
% The very first letter is a 2 line initial drop letter followed
% by the rest of the first word in caps.
% 
% form to use if the first word consists of a single letter:
% \IEEEPARstart{A}{demo} file is ....
% 
% form to use if you need the single drop letter followed by
% normal text (unknown if ever used by IEEE):
% \IEEEPARstart{A}{}demo file is ....
% 
% Some journals put the first two words in caps:
% \IEEEPARstart{T}{his demo} file is ....
% 
% Here we have the typical use of a "T" for an initial drop letter
% and "HIS" in caps to complete the first word.
\IEEEPARstart{T}{hroughout} this paper we denote by $f\vert_{\mathcal C} = \{y:\exists x\in\mathcal C\;y=f(x)\}$
the image of $\mathcal C$ under the map $f$, for polytopes defined by intersection of the halfspaces 
$a_i x \leq b_i$ for $i\in\{1,\dots,m\}$ we denote the matrix~${\bf{a}}_{\mathcal I}$ and vector~
${\bf{b}}_{\mathcal I}$ with the rows $a_i$ and $b_i$ for all indices in the index set $i\in\mathcal 
I\subseteq\{1,\dots,m\}$.

@@

  \mathcal W(p) := \left\{x: a_i x \leq \max_{k\leq l}\{b_{i,k} + c_{i,k}p\},i\leq m\right\}

@@

\item Notice that $h_i(p) = \max_{k\leq l} \{b_{i,k} + c_{i,k}p\}$ is a multi-parametric linear program (mpLP),
its solution is a piecewise affine function $h_i(p) = b_{i,k^\ast_i} + c_{i,k^\ast_i}p$ where $k^\ast_i(p)$ is constant

@@

  \mathscr G(\mathcal W) = \bigcup_{j\leq t} \left\{x: a_i x \leq b_{i,k_i^\ast} + c_{i,k_i^\ast}p,i\leq m

@@

$\left(b_{i,k_i^\ast} + c_{i,k_{j_1}^\ast} p\right)\big\vert_{\mathcal P_{j_1}} = 
\left(b_{i,k_{j_2}^\ast} + c_{i,k_{j_2}^\ast} p\right)\big\vert_{\mathcal P_{j_2}}$.

@@

  a_i v_\kappa(p) &= b_{i,k_i^\ast} + c_{i,k_i^\ast} p \quad\forall i\in\mathcal A_\kappa(p)\\
  a_i v_\kappa(p) &< b_{i,k_i^\ast} + c_{i,k_i^\ast} p \quad\forall i\in\mathcal I_\kappa(p)

@@

  \frac{d}{dp}\left(  {\bf{a}}_{\mathcal J\cup \{s\}} v_{\kappa_1}(p) - {\bf{b}}_{\mathcal J\cup \{s\}} - 
  {\bf{c}}_{\mathcal J\cup \{s\},{\bf{k}}^\ast} p\right) &= 0\\
  \frac{d}{dp}\left(  {\bf{a}}_{\mathcal J\cup \{u\}} v_{\kappa_2}(p) - {\bf{b}}_{\mathcal J\cup \{u\}} - 
  {\bf{c}}_{\mathcal J\cup \{u\},{\bf{k}}^\ast} p\right) &= 0

@@

  {\bf{a}}_{\mathcal J\cup \{s\}} \frac{dv_{\kappa}}{dp} &= {\bf{c}}_{\mathcal J\cup \{s\},{\bf{k}}^\ast}\\
  {\bf{c}}_{\mathcal J\cup \{u\}} \frac{dv_{\kappa}}{dp} &= {\bf{c}}_{\mathcal J\cup \{u\},{\bf{k}}^\ast}

@@

  \frac{dv_\kappa}{dp} = {\bf{a}}_{\mathcal J\cup \{s\}}^{-1}{\bf{c}}_{\mathcal J\cup \{s\},{\bf{k}}^\ast} = 
  {\bf{a}}_{\mathcal J\cup \{u\}}^{-1}{\bf{c}}_{\mathcal J\cup \{u\},{\bf{k}}^\ast}

@@

Corollary~\ref{thm:p:convexity:PWA:set:constant:num:verts} is essential for numerical evaluation, as it can be reformulated to:

@@

for a linear system~$x^+=\Psi x + v$ subject to disturbance $v\in\mathcal V(x)$ where $\mathcal V(x)$ is 
a piecewise affine parametrically convex point-to-set map defined 

@@

Since ${\mathcal{V}}(x)$ has a piecewise affine dependence on $x$ the vertices $v_i(x)$ are also piecewise 
affine in $x$.
The set $\mathcal X^\infty$ as defined in~\eqref{eq:definition:MRPI:set} is required to satisfy $\Psi x + 
v\in\mathcal X^\infty$ for all~$x\in\mathcal X^\infty$ and $v\in\mathcal V(x)$. 
To compute the MRPI set we start from the given state constraint set $\mathcal X_0=\mathcal X$

@@

and recursively \emph{cut off} points that cannot satisfy the invariance condition~\eqref{eq:definition:MRPI:set}

@@

$\mathcal X_0$. 
The first iteration enforces the constraint:

@@

Here $v_{0,i}^\ast(x)$ is not necessarily given by a unique maximiser, but is the solution of a multi-parametric 
linear program and hence is given by a vertex $v_i(x)$ of $\mathcal V(x)$ for all $x$ on that facet.

@@

@@

Determining the map $v_{l,i}(x)$ for all admissible $x\in \mathcal X$ in general poses a great classification

@@

Notice that we do not require $\Psi$ to be invertible, in~\eqref{eq:set:iteration:state:dependent:constraints} 
$\Psi^{-1}\mathcal X_k$ merely denotes the preimage of $\mathcal X_k$ under the linear map $\Psi$.

@@

Let the state constraint set $\mathcal X$ be contained in a band: $\mathcal X\subseteq B=\{x:\Gamma x\leq{\bf{1}}\wedge 
-\Gamma x\leq{\bf{1}}\}$, let the pair $(\Psi,\Gamma)$ be observable and let $\mathcal V(x)=D\mathcal W(x)$ with 
$\mathcal W(x)$ defined by~\eqref{eq:definition:PWA:polytopic:set:general}. 
Then $\mathcal X_N\subseteq \mathcal X_{N+1}$ for a finite $N$, and hence the MRPI set $\mathcal X^\infty =\mathcal X_N$ 
is a polytope.

@@

if $\mathcal X_k =\emptyset$ for any $k\geq 0$.
For the remainder of the proof we assume that the MRPI set $\mathcal X^\infty$ is non-empty. 
For this we require that $\bigcup_{s\in\mathcal S}\mathcal V(s)$ is bounded on any bounded set 
$\mathcal S\subset\mathbb R^n$.
The proof has two main steps: 
First we prove that $\mathcal X_p$ is bounded for $p\leq n$ where $n$ is the dimension of the state $x$. 
The second step is to prove that in~\eqref{eq:set:iteration:state:dependent:constraints}
the set $D_k$ grows exponentially, i.e. that for any given bounded set $\mathcal C$ there exists
finite $N$ such that $\mathcal C\subseteq D_{N}$. The proof is concluded by setting 
$\mathcal C = \mathcal X_{N}$ and deducing that $\mathcal X^\infty = \mathcal X_N$. 
For the first step, notice that observability of $(\Psi,\Gamma)$ is equivalent to the observability matrix 
$\Omega$ having full rank, i.e.\ $\mathrm{rank}(\Omega) = \mathrm{rank}(\begin{bmatrix} \Gamma^T \ \cdots \ 
(\Gamma\Psi^{n-1})^T\end{bmatrix}) = n$.
But  this implies that the set 
\mathcal P_{n} = \{x: 
\Omega x\leq{\bf{1}}\wedge-\Omega x\leq{\bf{1}}\} = \bigcap_{0\leq l\leq n-1} \Psi^{-l} B
is bounded, and since $\mathcal X_k\subseteq \mathcal P_k$ for all $k$, 
the set $\mathcal X_{n}$ is also bounded.

To see that $D_k$ grows exponentially we denote the spectral radius of $\Psi$ by $\rho$, since $\Psi$ is asymptotically
stable we have $\rho<1$.
Let $r_1$ be the radius of the largest ball contained in $\mathcal X_0$ i.e. $\mathcal E(r_1)\subseteq\mathcal X_0$ 
and let $r_2$ be the radius of smallest ball containing $\mathcal V(x)$ for all $x\in\mathcal X_0$.
Recall that $\mathcal X_k = \bigcap_{l\leq k} D_l$, where $D_k$ is defined by
  D_l = \underbrace{\Psi^{-l}\mathcal X_0}_{\mathcal S_1} \ominus 
  \underbrace{\left(\bigoplus_{1\leq i\leq l-1} \Psi^{i}\mathcal V(\mathcal X_{l-1})\right)}_{\mathcal S_2}
We can easily lower bound $\mathcal S_1$ by using $r_1$: $\mathcal S_1 = \Psi^{-l}\mathcal X_0 
\supseteq \Psi^{-l}\mathcal E(r_1) \supseteq \mathcal E(\rho^{-l} r_1)$.
Recall that $\mathcal X_l\subseteq\mathcal X_{l-1}$ and therefore $\mathcal X_l\subseteq\mathcal X_0$ for all $l\geq0$.
We use this fact to obtain the containment $\mathcal S_2 = \bigoplus_{1\leq i\leq l-1} 
\Psi^{i}\mathcal V(\mathcal X_{l-1})\subseteq \bigoplus_{1\leq i\leq l-1} \Psi^{i}\mathcal V(\mathcal X_0)\subseteq
\bigoplus_{1\leq i\leq l-1} \Psi^{i}\mathcal E(r_2)\subseteq\bigoplus_{1\leq i\leq l-1} \mathcal E(\rho^{i} r_2)\subseteq
\mathcal E(\sum_{1\leq i\leq l-1}\rho^i r_2)\subseteq \mathcal E(\frac{1}{1-\rho} r_2)$.
In summary we can state that~$D_l$ consists of the exponentially expanding set~$\mathcal S_1$
subtracted by the set~$\mathcal S_2$ which is bounded, therefore~$D_l$ itself expands exponentially,
that is $D_l\supseteq\mathcal E(\rho^{-l} r_1 - \frac{1}{1-\rho} r_2)$.
We have proven that for $p\leq n$ the set iterate~$\mathcal X_p$ is bounded, let $r_3$ denote the smallest 
ball that contains $\mathcal X_p$.
We can conclude that for $l$ such that $X_p\subseteq\mathcal E(r_3)\subseteq\mathcal 
E(\rho^{-l} r_1 - \frac{1}{1-\rho} r_2)\subseteq D_l$ holds the set $D_l$ covers
$X_p$, i.e. intersections with $D_l$ will not change $X_l$.
We can therefore give an upper bound on the number of iterations necessary for the algorithm
to terminate, it is given by the smallest integer $M$ such that
\begin{equation}\label{bnd:first:lower:bound:on:iteration:count}
  M\geq \frac{1}{\log(\frac{1}{\rho})}\left(\log\left(r_3+\frac{1}{1-\rho}r_2\right)-\log r_1 \right)
is satisfied.
Notice that the constants used in~\eqref{bnd:first:lower:bound:on:iteration:count} are not
convenient to compute, we can however compute bounds on $r_1,\, r_2$ and $r_3$ using 
singular values.
For this let $\bar\sigma(\Gamma)$ and $\underline\sigma(\Gamma)$ denote the maximal and the minimal
singular value of~$\Gamma$ respectively.
We know that~$\bar\sigma(\Gamma)\norm{x}\geq\norm{\Gamma x}$ and therefore $\bar\sigma(\Gamma)\norm{x}
\leq \sqrt{n}\Rightarrow \norm{\Gamma x}\leq\norm{\bf{1}}=\sqrt{n}$, that $r_1=\frac{\bar\sigma(\Gamma)}{\sqrt{n}}$
follows form standard result for singular values, see e.g.~\cite{Golub:1996}.
We use a similar argument to obtain a bound on $r_3$ before deriving a bound on $r_2$:
It is easy to see that the radius of the largest ball containing $\mathcal X_p$ can be determined using the maximal
norm of the vertices of $\mathcal P_n$, i.e. $r_3=\max_{i}\norm{x_i}$ where $x_i$ satisfies $\Omega_{\mathscr{A}_i} x_i={\bf{1}}$
and $\Omega_{\bar{\mathscr A}_i}x_i<{\bf{1}}$.
From this we can deduce that
\begin{multline}
  \sqrt{n}=\norm{\bf{1}}=\norm{\Omega_{\mathscr{A}_i}x_i}\geq\underline\sigma(\Omega_{\mathscr{A}_i})\norm{x_i}\\
  =\underline\sigma(\Gamma_{\mathscr{A}_i}\cdot \text{diag}(I,\Psi,\dots,\Psi^{n-1}))\norm{x_i}\\
  \geq\underline\sigma(\Gamma_{\mathscr{A}_i})\underline\sigma( \text{diag}(I,\Psi,\dots,\Psi^{n-1}))\norm{x_i}\\
  \geq\underline\sigma(\Gamma_{\mathscr{A}_i})\underline\sigma(\Psi)^{n-1}\norm{x_i}\\
  \geq\underline\sigma(\Gamma)\underline\sigma(\Psi)^{n-1}\norm{x_i},
\end{multline}
which implies that we can get the upper bound $r_3\leq\frac{\sqrt{n}}{\underline\sigma(\Gamma)
\underline\sigma(\Psi)^{n-1}}$.
In order to obtain an estimate on $r_2$ we follow a similar argumentation:
\begin{multline}
  \bar\sigma({\bf{c}})\norm{x}+\norm{{\bf{b}}}
  \geq\bar\sigma({\bf{c}}_{k^\ast})\norm{x}+\norm{{\bf{b}}_{k^\ast}}\\
  =\max_k\{\bar\sigma({\bf{c}}_k)\norm{x}+\norm{{\bf{b}}_k}\}
  \geq\norm{\max_k\{{\bf{c}}_{\mathscr{A}_i,k} x + {\bf{b}}_{{\mathscr{A}_i},k}\}}\\ 
  =\norm{{\bf{a}}_{\mathscr{A}_i}v_i} \geq\underline\sigma({\bf{a}}_{\mathscr A_i})\norm{v_i}
  \geq\underline\sigma({\bf{a}})\norm{v_i}.
\end{multline}
Notice that with only the assumptions made in Lemma~\ref{thm:finite:MRPI:set:state:dependable}
we have no way to bound~$\norm{x}$ in order to obtain an upper bound on~$r_2$,
however, if we assume that~$\mathcal X_p$ is not empty we can use the bound on~$\mathcal X_p$, i.e.~$r_3$.
This is reasonable since we are interested in the asymptotic behaviour of the set rather than
exact bounds on intermediate iterates, hence $r_2\leq\frac{\bar\sigma({\bf{c}})r_3+\norm{{\bf{b}}}}
{\underline\sigma({\bf{a}})}$ can be used as an estimate for the asymptotic behaviour of $\mathcal V(x)$
in the iteration.
All bounds and estimates on $r_1,\,r_2$ and $r_3$ can be calculated in advance to get an estimate
on how many iterations the described algorithm could require.
\section{Example I}\label{sec:example:I}
\begin{figure}
\centering
\begin{lpic}[scale=0.5]{levitatingBall}
\lbl[tr]{25,3; $m g$}
\lbl[br]{25,25; $c\frac{i^2}{y^2}$}
\lbl[bl]{49,17; $y$}
\lbl[bl]{56,55; $i$}
\end{lpic}
\vspace{-2mm}
\caption{Levitating ball system.}
\label{fig:levitating:ball}
\vspace{-2mm}\end{figure}
This section discusses the calculation of the MRPI set for a linearised, simplified model of the magnetic 
levitation system depicted in figure~\ref{fig:levitating:ball}. 
The system dynamics for the ball are given by $m \ddot y = m g - c\frac{i^2}{y^2}$, where $m,g,c,i$ and $y$ 
denote the mass of the ball, the gravitational constant, a constant factor, the current and the distance 
between the coil and the centre of the ball respectively.
For illustration purposes we neglect inductive dynamics, using the current $u=i$ as the control input and the position
$y$ and its first derivative $\dot y$ as the states: $x = (y,\dot y)^T$. We find that any equilibrium has 
$\dot{y}=0$ and $u=\sqrt{\frac{gm}{c}} y$ for any positive position $y>0$. 
Linearising the nonlinear differential equation $\dot x = f(x,u)$ around an equilibrium point $(\hat x, \hat
u)$ gives the approximate linear model 
   \Delta\dot{x} = \underbrace{\left(\begin{array}{cc}
  0 & 1 \\ \frac{2c\hat u^2}{m\hat x_1^3} & 0
  \end{array}\right)}_{\frac{\partial f}{\partial x}(\hat x,\hat
      u)}\Delta x 
+ \underbrace{\left(\begin{array}{c}
  0 \\ - \frac{2c\hat u}{m\hat x_1^2}
  \end{array}\right)}_{\frac{\partial f}{\partial u}(\hat x,\hat
      u)}\Delta u
where $\Delta u = u -\hat{u}$ and $\Delta x \approx x-\hat{x}$.
We derive the discrete time dynamics with sampling rate $T_s$ using the Euler formula $x^+=x+T_s f(x,u) 
=:\tilde f(x,u)$ giving $\Delta x^+ = A \Delta x + B \Delta u$,
A = I+T_s\frac{\partial f}{\partial  x}(\hat
  x,\hat u) , \quad
B = T_s \frac{\partial f}{\partial u}(\hat x,\hat u)
Although this system has a control input, the algorithm for computing the MRPI set is applicable since 
we consider the closed loop system under the linear feedback  $u=Kx$, where $K$ satisfies the aforementioned the robust 
Lyapunov condition $V(x)-V((A+BK)x+v)\leq \gamma^2v^Tv$ with $V(x)=x^T P x\geq0$, i.e. $x^TPx - ((A+BK)x+v)^TP(
(A+BK)x+v)\geq x^T(Q+K^TRK)x -\gamma^2 v^Tv$ for a minimum  $\gamma^2$, see e.g.~\cite{Boyd:94}.
A representation of additive disturbances acting on the linearised model due to linearisation errors, 
i.e. a system representation $x^+=Ax + Bu + v$ with an additive disturbance $v$, can be obtained using the
Mean Value Theorem (e.g.~\cite{Apostol:1974}).
\begin{thm}[Mean Value Theorem]\label{thm:mean:value:theorem}
Let $g :  X \rightarrow\mathbb R^m$ be continuously
differentiable, $ X\subset\mathbb R^n$ be open,
and $x \in X$, $h \in\mathbb R^n$ be such that 
$x + th \in X$ for all $t\in [0 ,1]$. Then
  g(x+h) = g(x) + \left(\int_0^1 \frac{\partial g}{\partial x}(x+th)dt\right)\cdot h.
Using the Mean Value Theorem and the linearisation around $(\hat{x},\hat{u})$, and defining 
$\tilde{x} = x - \hat{x}$, $\tilde{u} = u - \hat{u}$, we obtain the successor state $x^+ = 
\tilde{f}(\hat{x} + \tilde{x},\hat{u} + \tilde{u})$ as
\begin{multline*}
x^+= \tilde f(\hat x, \hat u) 
+ \int_0^1\frac{\partial\tilde f}{\partial x}(\hat x + t\tilde x,
\hat u + t\tilde u) dt \cdot \tilde x  \\
+ \int_0^1\frac{\partial\tilde f}{\partial u}(\hat x + t\tilde x,\hat u+
t\tilde{u}) dt\cdot \tilde{u}
\end{multline*}
and hence
\begin{multline*}
\tilde x^+ = A\tilde{x} + B \tilde{u} +\Bigl(
\int_0^1\frac{\partial\tilde f}{\partial x}(\hat x + t\tilde x,\hat u
             + t\tilde{u})dt - A \Bigr)\tilde x 
\\ 
+\Bigl(\int_0^1\frac{\partial\tilde f}{\partial u}(\hat x +
  t\tilde x,\hat u + t\tilde{u})dt - B\Bigr)\tilde{u}
\end{multline*}
This implies the dynamics
\tilde x^+ = A\tilde x+B\tilde{u} + H^x\tilde{x} + H^u \tilde{u}, 
where $H^x$ and $H^u$ can be determined by integration. 
If $f(x,u)$ is continuously differentiable for all $x\in X$ and $u\in U$, where
$X$, $U$ are compact sets, then we can approximate the values of $H^x \tilde{x}+ H^u \tilde{u}$ 
by a finite convex combination of extremal values, i.e. $H^x \tilde{x}+H^u \tilde{u}\in\conv_k
\{H^x_k \tilde{x} + H^u_k \tilde{u}\}$.
Thus we can introduce the element\-wise disturbance bound 
\begin{multline}\label{eq:definition:element:wise:constraints:on:nonlinearities}
\mathcal V(\tilde{x},\tilde{u})=\biggl\{v:\min_k\{
H^x_{k,i}\tilde{x}+H^u_{k,i}\tilde{u}\}\leq v_i\,\wedge 
\\ 
v_i \leq \max_k\{H^x_{k,i}\tilde{x}+H^u_{k,i}\tilde{u}\}, \, i =1,\dots,n\biggr\}.
\end{multline}
Notice that~\eqref{eq:definition:element:wise:constraints:on:nonlinearities} is of the form~\eqref{eq:definition:PWA:polytopic:set:general}
and the theory introduced earlier applies.
The additively perturbed linear system $\tilde x^+ = A\tilde x + B\tilde u + v$ accounts for all nonlinearities 
within $X\times U$ if $v\in\mathcal V(\tilde x,\tilde u)$. 
For general nonlinear systems finding the extremal values of $(H^x,H^u)$ is non-trivial.
To obtain values for $(H^x_k,H^u_k)$ we sample $X\times U$ and evaluate the integral 
expressions defining $(H^x_k,H^u_k)$ pointwise.
This leads to an inner approximation of the linearisation error set, however we can make it as tight as necessary
by increasing the number of samples.
For this we use the numerical values for the example of the levitating ball: $T_s=30\,\text{ms}$, $C=1$, 
$m=100\,\text{g}$, $\hat x_1 = 50\,\text{mm}$ and $\mathcal X=\{x:\abs{x_1- \hat x_1}\leq 1\,\text{mm}
\wedge \abs{x_2}\leq 105\,\text{mm}/\text{s}\}$, $\mathcal U=\{u:\abs{ u-\hat u}\leq10\,\text{mA}\}$.
Using a total of 25 samples for the computation of $(H^x,H^u)$ we obtain the invariant set shown 
in figure~\ref{fig:MRPI:set:levitating:ball}.
The algorithm for computing the MRPI set terminates after 3 iterations.
\begin{figure}
\centering
\begin{lpic}{invariantSetStateDependant(.65,)}
{\tiny
\lbl[r]{9,94; $0.1$}
\lbl[r]{9,86; $0.08$}
\lbl[r]{9,78; $0.06$}
\lbl[r]{9,70; $0.04$}
\lbl[r]{9,63; $0.02$}
\lbl[r]{9,55; $0$}
\lbl[r]{9,47; $-0.02$}
\lbl[r]{9,39; $-0.04$}
\lbl[r]{9,32; $-0.06$}
\lbl[r]{9,24; $-0.08$}
\lbl[r]{9,16; $-0.1$}
\lbl[t]{11,9; $-6$}
\lbl[t]{30,9; $-4$}
\lbl[t]{48,9; $-2$}
\lbl[t]{67.5,9; $0$}
\lbl[t]{86,9; $2$}
\lbl[t]{104,9; $4$}
\lbl[t]{123,9; $6$}
\lbl{120,3; $\times10^{-3}$}
}
{\small
\lbl{68,3; $\tilde x_1$}
\lbl{0,55,90; $\tilde x_2$}
}
\end{lpic}
\caption{The maximal robust positively invariant set for the levitating ball.}
\label{fig:MRPI:set:levitating:ball}
\vspace{-2mm}\end{figure}
\section{Recursive Multi-Parametric Quadratic Programming to Solve Min-Max Programs}\label{sec:recursive:mpQP:via:line:search}

In this section we discuss how multi-parametric quadratic programs (mpQP) can be solved efficiently with an active-set 
solver (see e.g.~\cite{Fletcher:1987}).
The active set of constraints will be determined by a line-search, we will also deal with \emph{degenerate
constraints}, which can arise during the line-search.
\subsection{Solving Recursive Equality Constraint mpQPs}\label{ssec:recursive:mpQP}
Assume we want to solve the mpQP
  J(\phi) = \left\{\begin{array}{rcl}
  \min_\theta& &\frac{1}{2} \theta^T R \theta + r^T\theta + V(\theta) \\
  \text{s.t.}& &C\theta\leq e+ T\phi
  \end{array}\right.
with the positive definite weight $R>0$ and the real parameter $\theta$. 
We furthermore assume that the constraints are such that the origin is an interior point for a vanishing parameter
$\phi=0$, i.e. $0\in\{\theta:C\theta< e\}$.
Using
  V(\theta) = \left\{\begin{split}
    \max_x & \;\frac{1}{2} x^T Q x + q^Tx\\
    \text{s.t.} &\; Ax  \leq b +S\theta
  \end{split}\right.
with $Q<0$ and the same assumption on the constraint set as for the minimisation.
We want to use an active-set solver to obtain the solution for $\phi=\phi^\ast$.
For this assume that for some $\phi=\phi_0$ we know the active constraints~$\mathcal A$ and~$\hat{\mathcal A}$,
i.e. the inequalities satisfied with equality $C_{\mathcal A}\theta_0=e_{\mathcal A} + T_{\mathcal A}\phi_0$
and $A_{\hat{\mathcal A}}x_0=b_{\hat{\mathcal A}}+S_{\hat{\mathcal A}}\theta_0$.
The set of inactive constraints $\mathcal I$ and $\hat{\mathcal I}$ are such that $C_{\mathcal I}\theta_0<
e_{\mathcal I} + T_{\mathcal I}\phi_0$ and $A_{\hat{\mathcal I}}x_0<b_{\hat{\mathcal I}}+S_{\hat{\mathcal I}}
\theta_0$, the variables~$\theta_0$ and $x_0$ denote the optimal solution to the minimisation and the maximisation
